Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-04 article-models-tt.tex
removed a false rule
typo
working on unfold
the unfold command works nicely now

\begin{defn}
For $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
We define
\begin{align*}
\jfamdefn*{{\Gamma}{A}}{\jcomp{f}{Q}}{\subst{f}{\ctxwk{A}{Q}}}\\
\jtermdefn*{{\Gamma}{A}}{\jcomp{f}{Q}}{\jcomp{f}{g}}{\subst{f}{\ctxwk{A}{g}}}.
\end{align*}
Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define
\begin{align*}
\jfamdefn*{{{\Gamma}{A}}{\jcomp{f}{Q}}}{\jcomp{f}{R}}{\subst{f}{\ctxwk{A}{R}}}\\
\jtermdefn*{{{\Gamma}{A}}{\jcomp{f}{Q}}}{\jcomp{f}{R}}{\jcomp{f}{h}}{\subst{f}{\ctxwk{A}{h}}}.
\end{align*}
\end{defn}

@@

\begin{lem}
We have the following inference rules about the situation where something is
substituted by a composition:
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}}
  {\jfameq
    {{\Gamma}{A}}
    {\subst{\jcomp{f}{g}}{R}}
    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}}
  {\jfameq
    {{\Gamma}{A}}
    {\subst{\jcomp{f}{g}}{h}}
    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
\end{align*}
We also have the following related inference rules, asserting that composition
is strictly associative:
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{\Gamma}{C}}{R}}
  {\jfameq
    {{\Gamma}{A}}
    {\jcomp{{f}{g}}{R}}
    {\jcomp{f}{{g}{R}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jterm{{\Gamma}{C}}{R}{h}}
  {\jtermeq
    {{\Gamma}{A}}
    {\jcomp{{f}{g}}{R}}
    {\jcomp{{f}{g}}{h}}
    {\jcomp{f}{{g}{h}}}}
\end{align*}
\end{lem}

\begin{proof}
Consider family morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
and a family $\jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}$. Then we have the judgmental
equalities
\begin{align*}
\subst{\jcomp{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}\\
& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
\end{align*}
The proof that 
$\subst{\jcomp{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
is similar.

Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have
\begin{align*}
\jcomp{{f}{g}}{R} & \jdeq \subst{\jcomp{f}{g}}{\ctxwk{A}{R}}\\
& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}\\
& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}\\
& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}\\
& \jdeq \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}\\
& \jdeq \subst{f}{\ctxwk{A}{\jcomp{g}{R}}}\\
& \jdeq \jcomp{f}{{g}{R}}.
\end{align*}
Again, the proof is similar for terms $h$ of $R$ in context $\ctxext{\Gamma}{C}$.
\end{proof}

\begin{lem}
We have the following inference rules about the compatibility of composition with
weakening:
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{\Gamma}{A}}{P}}
  {\jhomeq{\Gamma}{{A}{P}}{C}
    {\ctxwk{P}{\jcomp{f}{g}}}
    {\jcomp{\ctxwk{P}{f}}{g}}}\\
& \inference
  {\jterm{\Gamma}{B}{y}
   \jhom{\Gamma}{B}{C}{g}}
  {\jhomeq{\Gamma}{A}{C}
    {\jcomp{\ctxwk{A}{y}}{g}}
    {\ctxwk{A}{\subst{y}{g}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jterm{\Gamma}{C}{z}}
  {\jhomeq{\Gamma}{A}{C}
    {\jcomp{f}{\ctxwk{B}{z}}}
    {\ctxwk{A}{z}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{P}{\jcomp{f}{g}} 
& \jdeq \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}\\
& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}\\
& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}\\
& \jdeq \jcomp{\ctxwk{P}{f}}{g}.
\end{align*}
Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{\ctxwk{A}{y}}{g}
& \jdeq \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}\\
& \jdeq \ctxwk{A}{\subst{y}{g}}.
\end{align*}
For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
Then we have the judgmental equalities
\begin{align*}
\jcomp{f}{\ctxwk{B}{z}} 
& \jdeq \subst{f}{\ctxwk{A}{{B}{z}}}\\
& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}\\
& \jdeq \ctxwk{A}{z}.
\end{align*}
\end{proof}

\begin{lem}
We have the following inference rules about the compatibility of composition with
substitution:
\begin{align*}
& \inference
  {\jhom{{\Gamma}{A}}{P}{Q}{f}
   \jhom{{\Gamma}{A}}{Q}{R}{g}
   \jterm{\Gamma}{A}{x}}
  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}
    {\subst{x}{\jcomp{f}{g}}}
    {\jcomp{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jterm{\Gamma}{A}{x}}
  {\jtermeq{\Gamma}{C}
    {\subst{x}{\jcomp{f}{g}}}
    {\subst{{x}{f}}{g}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and $\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{f}{g}}
& \jdeq \subst{x}{{f}{\ctxwk{P}{g}}}\\
& \jdeq \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}\\
& \jdeq \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}\\
& \jdeq \jcomp{\subst{x}{f}}{\subst{x}{g}}.
\end{align*}
Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{f}{g}}
& \jdeq \subst{x}{{f}{\ctxwk{A}{g}}}\\
& \jdeq \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}\\
& \jdeq \subst{{x}{f}}{g}.
\end{align*}
\end{proof}

@@

\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. A morphism
from $P$ to $Q$ over $f$ in context $\Gamma$ is a term
\begin{equation*}
\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}.
\end{equation*}
\end{defn}

@@

\subsubsection{Identity functions}
Without a rule explicitly asserting the existence of an identity morphism we don't
get one, hence we do that here. The identity morphism is a term which introduced
in ordinary type theory via the variable rule. The variable rule is a bit more
general: it asserts that
\begin{equation*}
\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
\end{equation*}
for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
we get the projections from the identity morphisms together with weakening. We
already have weakening, so here it suffices to introduce the identity morphisms.
\begin{align}
& \inference
  {\jfam{\Gamma}{A}}
  {\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}}
& & \inference
    {\jfameq{\Gamma}{A}{A'}}
    {\jtermeq{{\Gamma}{A}}{\ctxwk{A}{A}}{\idfunc[A]}{\idfunc[A']}}
\end{align}
Identity functions are determined by their behavior with respect to substitution combined with
weakening. The identity functions will also be subject to compatibility rules.
\begin{align}
& \inference
  {\jterm{\Gamma}{A}{x}}
  {\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}
  \label{idfunc-subst-defn}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}
  \label{idfunc-wk-defn}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}
  \label{idfunc-wk-defn2}\\
& \inference
  {\jterm{{\Gamma}{A}}{P}{f}}
  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}
  \label{idfunc-precomp}\\
& \inference
  {\jterm{{\Gamma}{A}}{P}{f}}
  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}
  \label{idfunc-precomp}\\
& \inference
  {\jterm{{\Gamma}{B}}{\ctxwk{B}{A}}{g}}
  {\jtermeq{{\Gamma}{B}}{\ctxwk{B}{A}}{\subst{g}{\ctxwk{B}{\idfunc[A]}}}{g}}
  \label{idfunc-postcomp}
\end{align}

@@

The identity function of a weakened family is the weakened identity function:
\begin{equation}\label{idfunc-wk-comp}
\inference
  {\jfam{\Gamma}{A}
   \jfam{\Gamma}{B}}
  {\jtermeq{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{B}}}{\idfunc[\ctxwk{A}{B}]}{\ctxwk{A}{\idfunc[B]}}}
\end{equation}

The identity function of a substituted family is the substitution of the identity function
\begin{equation}\label{idfunc-subst-comp}
\inference
  {\jterm{\Gamma}{A}{x}
   \jfam{{\Gamma}{A}}{P}}
  {\jtermeq{{\Gamma}{\subst{x}{P}}}{\ctxwk{\subst{x}{P}}{\subst{x}{P}}}{\idfunc[\subst{x}{P}]}{\subst{x}{\idfunc[P]}}}
\end{equation}

@@

\begin{lem}
The following inference rule is valid
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{\Gamma}{C}}{R}
   \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}}
  {\jhomeq{\Gamma}{A}{{C}{R}}
    {\jcomp{f}{\ctxext{g}{w}}}
    {\ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{f}{\ctxext{g}{w}}
& \jdeq \subst{f}{\ctxwk{A}{\ctxext{g}{w}}}\\
& \jdeq \subst{f}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
& \jdeq \ctxext{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
& \jdeq \ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}.
\end{align*}
\end{proof}
%\end{document}

@@

\subsubsection{Families of extensions}
\begin{equation*}
\inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfameq{{{\Gamma}{A}}{P}}
    {\mfam{{A}{P}}}
    {{\ctxwk{P}{\mfam{A}}}{\ctxwk{{P}{\mfam{A}}}{\mfam{P}}}}}
\end{equation*}

>>>>

\begin{defn}
We define the judgment
\begin{equation*}
\jhom{\Gamma}{A}{B}{f},
\end{equation*}
which is pronounced as `$f$ is a morphism from $A$ to $B$ in context $\Gamma$',
to be the judgment
\begin{equation*}
\unfold{\jhom{\Gamma}{A}{B}{f}}.
\end{equation*}
Likewise, we define the judgment
\begin{equation*}
\jhomeq{\Gamma}{A}{B}{f}{f'}
\end{equation*}
to be the judgment
\begin{equation*}
\unfold{\jhomeq{\Gamma}{A}{B}{f}{f'}}.
\end{equation*}
\end{defn}

\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ and consider a term $\jterm{{\Gamma}{B}}{Q}{g}$.
We define
\begin{align*}
\jfamdefn*
  {{\Gamma}{A}}
  {\jcomp{A}{f}{Q}}
  {\unfold{\jcomp{A}{f}{Q}}}\\
\jtermdefn*
  {{\Gamma}{A}}
  {\jcomp{A}{f}{Q}}
  {\jcomp{A}{f}{g}}
  {\unfold{\jcomp{A}{f}{g}}}.
\end{align*}
Likewise, when we have a family $\jfam{{{\Gamma}{B}}{Q}}{R}$ and a term
$\jterm{{{\Gamma}{B}}{Q}}{R}{h}$, we define
\begin{align*}
\jfamdefn*{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}{\unfold{\jcomp{A}{f}{R}}}\\
\jtermdefn*{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}{\jcomp{A}{f}{h}}{\unfold{\jcomp{A}{f}{h}}}.
\end{align*}
\end{defn}

@@

\begin{lem}
We have the following inference rules about the situation where something is
substituted by a composition:
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
    \jhom{\Gamma}{B}{C}{g}
    \jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}}
  {\jfameq
    {{\Gamma}{A}}
    {\subst{\jcomp{A}{f}{g}}{R}}
    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
    \jhom{\Gamma}{B}{C}{g}
    \jterm{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}{h}}
  {\jfameq
    {{\Gamma}{A}}
    {\subst{\jcomp{A}{f}{g}}{h}}
    {\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}}}
\end{align*}
We also have the following related inference rules, asserting that composition
is strictly associative:
\begin{align*}
& \inference
    {\jhom{\Gamma}{A}{B}{f}
      \jhom{\Gamma}{B}{C}{g}
      \jfam{{\Gamma}{C}}{R}}
    {\jfameq
      {{\Gamma}{A}}
      {\jcomp{A}{{A}{f}{g}}{R}}
      {\jcomp{A}{f}{{B}{g}{R}}}}\\
& \inference
    {\jhom{\Gamma}{A}{B}{f}
      \jhom{\Gamma}{B}{C}{g}
      \jterm{{\Gamma}{C}}{R}{h}}
    {\jtermeq
      {{\Gamma}{A}}
      {\jcomp{A}{{A}{f}{g}}{R}}
      {\jcomp{A}{{A}{f}{g}}{h}}
      {\jcomp{A}{f}{{B}{g}{h}}}}
\end{align*}
\end{lem}

\begin{proof}
Consider family morphisms $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{\Gamma}{B}{C}{g}$
and a family $\jfam{{{\Gamma}{A}}{\ctxwk{A}{C}}}{R}$. Then we have the judgmental
equalities
\begin{align*}
\subst{\jcomp{A}{f}{g}}{R} & \jdeq \subst{{f}{\ctxwk{A}{g}}}{R}\\
& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{{A}{B}}{R}}}\\
& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{R}}}
\end{align*}
The proof that 
$\subst{\jcomp{A}{f}{g}}{h}\jdeq\subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{h}}}$
is similar.

Now suppose that $\jfam{{\Gamma}{C}}{R}$ instead. Then we have
\begin{align*}
\jcomp{A}{{A}{f}{g}}{R} & \jdeq \subst{\jcomp{A}{f}{g}}{\ctxwk{A}{R}}\\
& \jdeq \subst{{f}{\ctxwk{A}{g}}}{\ctxwk{A}{R}}\\
& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{{A}{B}}{{A}{R}}}}\\
& \jdeq \subst{f}{{\ctxwk{A}{g}}{\ctxwk{A}{{B}{R}}}}\\
& \jdeq \subst{f}{\ctxwk{A}{\subst{g}{\ctxwk{B}{R}}}}\\
& \jdeq \subst{f}{\ctxwk{A}{\jcomp{B}{g}{R}}}\\
& \jdeq \jcomp{A}{f}{{B}{g}{R}}.
\end{align*}
Again, the proof is similar for terms $h$ of $R$ in context $\ctxext{\Gamma}{C}$.
\end{proof}

\begin{lem}
We have the following inference rules about the compatibility of composition with
weakening:
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{\Gamma}{A}}{P}}
  {\jhomeq{\Gamma}{{A}{P}}{C}
    {\ctxwk{P}{\jcomp{A}{f}{g}}}
    {\jcomp{{A}{P}}{\ctxwk{P}{f}}{g}}}\\
& \inference
  {\jterm{\Gamma}{B}{y}
   \jhom{\Gamma}{B}{C}{g}}
  {\jhomeq{\Gamma}{A}{C}
    {\jcomp{A}{\ctxwk{A}{y}}{g}}
    {\ctxwk{A}{\subst{y}{g}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jterm{\Gamma}{C}{z}}
  {\jhomeq{\Gamma}{A}{C}
    {\jcomp{A}{f}{\ctxwk{B}{z}}}
    {\ctxwk{A}{z}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jfam{{\Gamma}{A}}{P}$.
Then we have the judgmental equalities
\begin{align*}
\ctxwk{P}{\jcomp{A}{f}{g}} 
& \jdeq \ctxwk{P}{\subst{f}{\ctxwk{A}{g}}}\\
& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{P}{{A}{g}}}\\
& \jdeq \subst{\ctxwk{P}{f}}{\ctxwk{\ctxext{A}{P}}{g}}\\
& \jdeq \jcomp{{A}{P}}{\ctxwk{P}{f}}{g}.
\end{align*}
Now let $\jterm{\Gamma}{B}{y}$ and $\jhom{\Gamma}{B}{C}{g}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{\ctxwk{A}{y}}{g}
& \jdeq \subst{\ctxwk{A}{y}}{\ctxwk{A}{g}}\\
& \jdeq \ctxwk{A}{\subst{y}{g}}.
\end{align*}
For the third assertion, let $\jhom{\Gamma}{A}{B}{f}$ and $\jterm{\Gamma}{C}{z}$.
Then we have the judgmental equalities
\begin{align*}
\jcomp{A}{f}{\ctxwk{B}{z}} 
& \jdeq \subst{f}{\ctxwk{A}{{B}{z}}}\\
& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{z}}}\\
& \jdeq \ctxwk{A}{z}.
\end{align*}
\end{proof}

\begin{lem}
We have the following inference rules about the compatibility of composition with
substitution:
\begin{align*}
& \inference
  {\jhom{{\Gamma}{A}}{P}{Q}{f}
   \jhom{{\Gamma}{A}}{Q}{R}{g}
   \jterm{\Gamma}{A}{x}}
  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{R}}
    {\subst{x}{\jcomp{P}{f}{g}}}
    {\jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jterm{\Gamma}{A}{x}}
  {\jtermeq{\Gamma}{C}
    {\subst{x}{\jcomp{A}{f}{g}}}
    {\subst{{x}{f}}{g}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{{\Gamma}{A}}{P}{Q}{f}$, $\jhom{{\Gamma}{A}}{Q}{R}{g}$ and $\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{A}{f}{g}}
& \jdeq \subst{x}{{f}{\ctxwk{P}{g}}}\\
& \jdeq \subst{{x}{f}}{{x}{\ctxwk{P}{g}}}\\
& \jdeq \subst{{x}{f}}{\ctxwk{\subst{x}{P}}{\subst{x}{g}}}\\
& \jdeq \jcomp{\subst{x}{P}}{\subst{x}{f}}{\subst{x}{g}}.
\end{align*}
Now let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$ and $\jterm{\Gamma}{A}{x}$.
Then we have the judgmental equalities
\begin{align*}
\subst{x}{\jcomp{A}{f}{g}}
& \jdeq \subst{x}{{f}{\ctxwk{A}{g}}}\\
& \jdeq \subst{{x}{f}}{{x}{\ctxwk{A}{g}}}\\
& \jdeq \subst{{x}{f}}{g}.
\end{align*}
\end{proof}

@@

\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism from $A$ to $B$ in context $\Gamma$
and consider $\jfam{{\Gamma}{A}}{P}$ and $\jfam{{\Gamma}{B}}{Q}$. We define the
judgment
\begin{equation*}
\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F},
\end{equation*}
which is pronounced as `$F$ is a morphism from $P$ to $Q$ over $f$ in context
$\Gamma$, to be the judgment
\begin{equation*}
\unfold{\jfhom{\Gamma}{A}{B}{f}{P}{Q}{F}}.
\end{equation*}
\end{defn}

@@

\subsubsection{Identity functions}
Without a rule explicitly asserting the existence of an identity morphism we don't
get one, hence we do that here. The identity morphism is a term which introduced
in ordinary type theory via the variable rule. The variable rule is a bit more
general: it asserts that
\begin{equation*}
\jterm{\Gamma,\,x_1:A_1,\ldots,\,x_n:A_n}{A_i}{x_i}
\end{equation*}
for every $1\leq i\leq n$. Thus, it establishes the projections. In our setting,
we get the projections from the identity morphisms together with weakening. We
already have weakening, so here it suffices to introduce the identity morphisms.
\begin{align}
& \inference
  {\jfam{\Gamma}{A}}
  {\jhom{\Gamma}{A}{A}{\idfunc[A]}}
& & \inference
    {\jfameq{\Gamma}{A}{A'}}
    {\jhomeq{\Gamma}{A}{A}{\idfunc[A]}{\idfunc[A']}}
\end{align}
Identity functions are determined by their behavior with respect to substitution combined with
weakening. The identity functions will also be subject to compatibility rules.
\begin{align}
& \inference
  {\jterm{\Gamma}{A}{x}}
  {\jtermeq{\Gamma}{A}{\subst{x}{\idfunc[A]}}{x}}
  \label{idfunc-subst-defn}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}{P}}
  \label{idfunc-wk-defn}\\
& \inference
  {\jfam{{\Gamma}{A}}{P}}
  {\jfameq{{\Gamma}{A}}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{P}}}{P}}
  \label{idfunc-wk-defn2}\\
& \inference
  {\jterm{{\Gamma}{A}}{P}{f}}
  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{A}{f}}}{f}}
  \label{idfunc-precomp}\\
& \inference
  {\jterm{{\Gamma}{A}}{P}{f}}
  {\jtermeq{{\Gamma}{A}}{P}{\subst{\idfunc[A]}{\ctxwk{{A}{A}}{f}}}{f}}
  \label{idfunc-precomp}\\
& \inference
  {\jhom{\Gamma}{A}{B}{f}}
  {\jhomeq{\Gamma}{A}{B}{\jcomp{A}{f}{\idfunc[B]}}{f}}
  \label{idfunc-postcomp}
\end{align}

@@

The identity function of a weakened family is the weakened identity function:
\begin{equation}\label{idfunc-wk-comp}
\inference
  {\jfam{\Gamma}{A}
   \jfam{\Gamma}{B}}
  {\jhomeq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A}{B}}{\ctxwk{A}{\idfunc[B]}}{\idfunc[\ctxwk{A}{B}]}}
\end{equation}

The identity function of a substituted family is the substitution of the identity function
\begin{equation}\label{idfunc-subst-comp}
\inference
  {\jterm{\Gamma}{A}{x}
   \jfam{{\Gamma}{A}}{P}}
  {\jhomeq{\Gamma}{\subst{x}{P}}{\subst{x}{P}}{\subst{x}{\idfunc[P]}}{\idfunc[\subst{x}{P}]}}
\end{equation}

@@

\begin{lem}
The following inference rule is valid
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{\Gamma}{C}}{R}
   \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}}
  {\jhomeq{\Gamma}{A}{{C}{R}}
    {\jcomp{A}{f}{\ctxext{g}{w}}}
    {\ctxext{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{A}{f}{\ctxext{g}{w}}
& \jdeq \subst{f}{\ctxwk{A}{\ctxext{g}{w}}}\\
& \jdeq \subst{f}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
& \jdeq \ctxext{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
& \jdeq \ctxext{\jcomp{A}{f}{g}}{\jcomp{A}{f}{w}}.
\end{align*}
\end{proof}
%\end{document}

@@

\subsubsection{Families of extensions}
